---
title: "BST 260 Final Project"
author: "Tiffany Zeng"
subtitle: Factors Associated with Healthcare Utilization Among Elderly in Shanghai,
  China
output: html_document
---
---
title: "bst260finalproject"
output: github_document
date: "2022-12-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction 

China’s growing aging population can be seen as a threat to the country’s healthcare system by pushing the government to shift their resources to address the needs of the older population, therefore putting more strain on the healthcare system. Although life expectancy has increased by 17 years over the past 2 decades, the older population experiences more years lived with disability due to chronic conditions (Kennedy et al., 2014). The main challenge with a large elderly population is that because almost 80% of deaths in China are attributed to chronic NCDs among those aged 60 years and older, therefore it is important to understand the factors that drive healthcare utilization among China's aging population.

According to the IHME, out of the top 10 causes of death for those ages 70 and older, 9 are NCDs, with cardiovascular diseases, neoplasms, and chronic respiratory diseases to be the top 3 causes of death (Vos et al., 2020). Additionally, with the growing older population, China will be losing 70 million of its workforce over the next few years, which will be a huge shock to the global supply chain as much of the world depends on China for its manufacturing factories.By ensuring a healthier elderly population, more people would remain in the workforce and therefore continue to contribute to China’s growing economy. There is a saying that China will “get old before it gets rich,” which accurately describes what the country is currently experiencing. A growing elderly population coupled with a shrinking pool of working adults will add pressure to the country’s health and social welfare system. 

The objective of this study is to focus on the health status of Shanghai's aging population, as Shanghai has shown a more dramatic growth in its aging population in recent years. According to the Shanghai Bureau of Statistics, the city had a population of 14.5 million people living in households by the end of 2016, of which over 31% were aged 60 and over (Shanghai Bureau of Statistics, 2011).

The methodology of identifying this dataset was through datasetsearch.research.google.com and entering keywords "healthcare utilization among elderly population in China." The dataset used in this study is publicly available and sourced from a cross-sectional study that developed a questionnaire used for face-to-face interviews conducted in June and August 2011 among Shanghai's 18 districts.  

### Results

To begin the analysis, we are going to load the dataset: 

#### Load Dataset

```{r}
library(haven)
library(readxl)
library(tidyverse)
library(dplyr)
library(mlogit)

data <- read_excel("elderly_china_data.xlsx")
head(data)

```
We see that some of the variables are either spelled or labeled incorrectly, therefore we need to recode these variables before beginning our analysis. 

#### Rename Variables
```{r}
colnames(data)[8] = "insurance"
colnames(data)[9] = "pension_income"
colnames(data)[10] = "income_source"
colnames(data)[15] = "nervous"
colnames(data)[16] = "life_satisfaction"
colnames(data)[18] = "physical_health_change"
colnames(data)[23] = "Heart Disease" 
colnames(data)[28] = "regional_economic_level"
colnames(data)[29] = "outdoor_activity"
colnames(data)[31] = "community_activities"
colnames(data)[32] = "seeing_children"
colnames(data)[36] = "hospitalization"
colnames(data)[37] = "total_chronic"
```

We also want to double-check if there are any classification discrepancies in the dataset as well as generate variables of interest by converting them to binary variables if necessary. We are creating a binary "chronic" variable to label those in our study population that are diagnosed with 1 or more chronic conditions as "1" and those who do not have chronic conditions as "0". Additionally, we are creating a "married" and "never_married" variable to further disaggregate the "marry" variable. Lastly, we are generating a second gender and age variable to easily identify their categories.

#### Convert Variable Types/Generate Variables
```{r}
class(data$gender)

data$gender2 <- factor(ifelse(data$gender == 1, "Male", "Female"))
data <- data |>
  mutate(age2 = factor (
    x = age,
    levels = 1:5,
    labels = c("60_64", "65_69", "70_74", "75_79", ">=80"
  )))

#generate binary chronic variable 
data$chronic <- ifelse(data$total_chronic>=1, 1, 0)

#generate indicator variables for married and never married
data$married <- ifelse(data$marry == 3, 1, 0)
data$never_married <- ifelse(data$marry == 1 & 2, 1, 0)
```
#### Summarize Variables
```{r}
data |>
  group_by(gender) |>
  summarize(n())

data |>
  group_by(age) |>
  summarize(n())

data |>
  group_by(race) |>
  summarize(n())

data |>
  group_by(education) |>
  summarize(n())

data |>
  group_by(children) |>
  summarize(n())
```

Lastly, summarizing variables of interest gives us a better understanding of the percentage breakdown within each category. Extensive data wrangling was necessary for this dataset since we wanted to manipulate variables that are specific to this analysis. The original dataset had many errors that would not have allowed for appropriate analyses. 

#### Linear Regression 
```{r}
m1 <- lm(chronic ~ gender + age + married + never_married + race + education + children + livingstatus + insurance + pension_income + location + lonely + smoking + drink, 
         data=data)

summary(m1)
```
The first model we are interested in building is a simple linear regression. We see that the "location" variable is significant at the 0.1 level, the "age" and "insurance" variables are significant at the 0.05 level, and the "lonely" variable is significant at the 0.0001 level. 

#### Logistic Regression
```{r}
m2 <- glm(chronic ~ gender + age + married + never_married + race + education + children + livingstatus + insurance + pension_income + location + lonely + smoking + drink, 
          family = binomial(link = "logit"),
               data = data)

summary(m2)
```
Because a simple linear regression is oftentimes not a feasible method to measure the effects of different factors on chronic health outcomes, a better method would be a logistic regression. In this analysis, we also see that the "location" variable is significant at the 0.1 level, the "age" and "insurance" variables are significant at the 0.05 level, and the "lonely" variable is significant at the 0.0001 level. 

#### Poisson Test 
```{r}
m3 <- glm(chronic ~ gender + age + married + never_married + race + education + children + livingstatus + insurance + pension_income + location + lonely + smoking + drink, 
          family = "poisson",
               data = data)

summary(m3)

library(mfx)
library(sandwich)
library(lmtest)
library(zoo)

poissonmfx(formula = chronic ~ gender + age + married + never_married + race + education + children + livingstatus + insurance + pension_income + location + lonely + smoking + drink, 
           data=data)
```
Although a Poisson model would not be too informative for this analysis, I wanted to explore whether the "lonely" variable was still significant in this context, which it is. 

#### Chi-squared Test
```{r}
chisq <- chisq.test(data$chronic, data$lonely)
chisq
```
Because the "lonely" variable has consistently shown to be significant, I wanted to do a chi-squared test between "chronic" and "lonely" to validate that an elderly person experiencing loneliness is a significant indicator to whether they are diagnosed by chronic conditions. Because p < 0.05, it is in fact significant. 

### Conclusion 
The research question of this study is to explore the effect of different factors on developing chronic conditions among the elderly population in China. The results showed the impact of economic status, health status, demographic and social characteristics on developing chronic conditions. The most surprising finding was that loneliness was a significant indicator of whether an elderly person develops chronic conditions, pointing towards the shift in Chinese culture where traditionally, children would eventually become the main caretakers of their parents. However, the Only-Child Policy in China has impacted this tradition as parents increasingly encourage their children to pursue their careers and not worry about their health, causing many children to not return to their families as often as they would if each family was still able to raise more children. This policy has proven to be devastating and burdening on China's health system as doctors are pushed to see more patients. 